Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-12-03 emu_JHEP_REsubmission/responseToEditor.txt

Dear Editor,

We thank the referee for their insightful feedback, agree with the referee that the manuscript is an excellent fit for JHEP, and have made suggested corrections.

@@

We have addressed all the minor clarifications:

C1] All appearances of 'Gramm determinants' have been changed to 'Gram determinants', and the appearance of 'veirbein' has been corrected to 'vierbein'.
C2] In 2.49) 'C' has been corrected to 'C_3'
C3] In 3.14 and 3.31, we have corrected 'l^{\mu_2}' to the uncontracted 'l^{\mu_3}'
C4] Below 4.39 we have corrected the doubled word 'use use'
C5] We have changed 'present as tree-level' to 'present at tree-level'
C6] At the referees request, we have changed ' adjoint double' to 'adjoint double copy' three lines about section 5.2.2

Furthermore, we very much appreciative the thoughtful and substantive feedback from the referee, and have made the following additions to the manuscript and provide the following clarification,

Q1] We agree that two-loop four-point is a particularly simple example that permits constructibility in terms of one-loop integral convolutions. Indeed this is precisely the reason that we focus on these amplitudes as a sharp example. We have clarified by changing

@@

Q3] We appreciate the referee's willingness to maintain the ostrich integrals as drawn, even though we agree that it obscures the Z2 symmetry of the diagram. We have added a number a couple reminders that the integral itself is symmetric under (1<->2) exchange. First in figure 3:

"Throughout the text, ostrich integrals will be drawn with an asymmetry between the “head” and the “tail” of the ostrich, but we emphasize that the integral itself is functionally symmetric under exchange of the top two external edges."

Then below eq. 4.23:

"As noted below Fig. 3, the ostrich integral above is symmetric under 1 <-> 2 exchange of external edges."

@@

"However, we should emphasize that eq. (4.38) only considers the leading divergence at each loop order, and does not consider subleading logarithmic terms related to the cross terms of (A1-loop/Atree)2 ∼ ε−2 + ε−1log(μ) + O(1). We leave determining whether eq. (4.38) also captures the subleading divergences as a direction of future study."

Q5] We have spelled out the result of plugging in the condition for maximal supersymmetry in D=4 -- these can be found in eq. 4.67 in D-dimensions and eq. 4.72 for D=4

Furthermore, we have emphasized that the particle content of maximal SUSY in 4.82 -- 4.85 is explicitly for a single supersymmetric vector multiplet, such that $N_\gamma = 1$):

''When one plugs in the state content for a single maximally supersymmetric vector multiplet (i.e. with normalization N_\gamma = 1), stated below for a variety of dimensions''

@@

Q6] Below eq. 5.47 we provide some background on how to derive these Hilbert series. Finding generating functions in general can be challenging, but since both eq 5.46 and 5.47 iteratively compound, we they can be easily derived from appropriate multiplication of 1/(1-\alpha). In eqs. 5.48 -- 5.51 we give an explicit example of how one can derive the permutation invariant Hilbert series computed in 5.46.

In addition, we have fixed the following typos that we identified in a final read through:

### 2023-12-03 emu_JHEP_REsubmission/emu.tex
minor corrections

For example, consider the two-loop divergence for the $(++++)$ anomalous matrix element in pure BI theory. The 4D helicity structure can be captured by a symmetric structure double-copy between NLSM pion numerators and a local gauge theory contact at $\mathcal{O}(\alpha'^5)$. The $s$-channel numerators for this symmetric-structure double copy are as follows
\begin{equation}\label{eq:symNum2loop}
n_s^{\text{dd,}\pi} = s^2 \qquad n_s^{\text{dd,HD}} = s^5 \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
Double-copying these kinematic numerators yields a matrix element of the form:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{BI}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{n_g^{\text{dd,HD}}n^{\text{dd,}\pi}_g}{d_g} = (s^6+t^6+u^6) \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
While this construction lacks any algebraic relations between the four-point kinematic-factors, similar symmetric numerators were found at six-point for NLSM, which obey non-trivial algebraic relations \cite{Carrasco:2022jxn}. In addition to the photon counterterms constructed above, symmetric double copy is likewise a natural description of gravitational counterterms. Rather than composing the local vector numerators, $n^{\text{dd,HD}}$, with symmetric NLSM numerators, $n^{\text{dd,}\pi}_g$, they can also be squared -- yielding a gravitational counterterm:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{GR}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{(n^{\text{dd,HD}}_g)^2}{d_g} = (s^9+t^9+u^9)  \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
In general, any four-point symmetric vector numerator \cite{Carrasco:2022jxn} can be constructed from two $D$-dimensional gauge theory building blocks,
\begin{equation}
n_s^{\text{dd,vec,1}} = f_{12}f_{34}\,, \qquad n_s^{\text{dd,vec,2}} = f_{1324}\,,
\end{equation}
by composing with the symmetric scalar numerators, $n_s^{\text{dd,1}} = s$ and  $n_s^{\text{dd,2}} = tu$. For example, with these building blocks, symmetric double-copy construction captures the exceptional four graviton amplitude of Ref.~\cite{Boels:2016xhc}, which only considered linear combinations of gauge theory amplitudes rather than gauge theory numerators. We note that a similar construction was recently identified in Ref.~\cite{Bourjaily:2023tcc} beyond four-point, which demonstrated that gravity amplitudes permit double copy construction in terms of gauge-invariant numerators that do not obey any algebraic relations between themselves, much like the symmetric numerators above in \eqn{eq:symNum2loop}. We see further exploring higher multiplicity examples of symmetric double copy as an important direction of future study.

@@

[\mathcal{H}^{(ijkl)}] = 1,0,1,1,1,1,2,1,2,2,2,2,3,2,3,3,3,...

@@

[\mathcal{H}^{(ij)(kl)}] = 1,1,2,2,3,3,4,4,5,5,6,6,7,7,8,8,...

@@

Since each of these Hilbert series has an iterative structure where they compound at successively higher orders, they can be easily derived. Take for example the Hilbert series for symmetric invariants, which grows periodically at $\mathcal{O}(\alpha^{6n})$. To derive this series, we can with the geometric series,

@@

\left[\frac{1}{1-\alpha} \right]= 1,1,1,1,1,1,1,1,\dots 

@@

\left[\frac{1-\alpha+\alpha^2}{1-\alpha} \right]= 1,0,1,1,1,1,1,1,\dots 

@@

\mathcal{H}^{(ijkl)} =  \frac{1-\alpha+\alpha^2}{1-\alpha} (1+\alpha^{6}+\alpha^{12}+\alpha^{18}+\cdots )

@@

\mathcal{H}^{(ijkl)} = \frac{1-\alpha+\alpha^2}{1-\alpha} \frac{1}{1-\alpha^6} =  \frac{1}{(\alpha-1)^2(\alpha+1)(\alpha^2+\alpha+1)} 

@@

\mathcal{T}^{F^2F^2}_{(x,y)} \sim s_{ij}^x(s_{ik}s_{jk})^y \qquad \mathcal{T}^{F^4}_{(x,y)} \sim s_{ij}^x(s_{ik}s_{jk})^y \qquad \mathcal{T}^{F^3}_{(x,y)} \sim \sigma_3^x\sigma_2^y\,,

@@

\mathcal{T}^{(--++)}_{(x,y)} \sim s_{ij}^x(s_{ik}s_{jk})^y \qquad \mathcal{T}^{(-+++)}_{(x,y)} \sim \sigma_3^x\sigma_2^y\qquad \mathcal{T}^{(++++)}_{(x,y)} \sim \sigma_3^x\sigma_2^y\,.

>>>>

For example, consider the two-loop divergence for the $(++++)$ anomalous matrix element in pure BI theory. The 4D helicity structure can be captured by a symmetric structure double-copy between NLSM pion numerators and a local gauge theory contact at $\mathcal{O}(\alpha'^5)$. The $s$-channel numerators for this symmetric-structure double copy are as follows,
\begin{equation}\label{eq:symNum2loop}
n_s^{\text{dd,}\pi} = s^2 \,, \qquad n_s^{\text{dd,HD}} = s^5 \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
Double-copying these kinematic numerators yields a matrix element of the form:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{BI}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{n_g^{\text{dd,HD}}n^{\text{dd,}\pi}_g}{d_g} = (s^6+t^6+u^6) \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
While this construction lacks any algebraic relations between the four-point kinematic-factors, similar symmetric numerators were found at six-point for NLSM, which obey non-trivial algebraic relations \cite{Carrasco:2022jxn}. In addition to the photon counterterms constructed above, symmetric double copy is likewise a natural description of gravitational counterterms. Rather than composing the local vector numerators, $n^{\text{dd,HD}}$, with symmetric NLSM numerators, $n^{\text{dd,}\pi}_g$, they can also be squared -- yielding a gravitational counterterm:
\begin{equation} \label{eq:2loopCTsymDC}
\mathcal{M}^{\text{GR}+\text{HD}} = \sum_{g\in \Gamma^3} \frac{(n^{\text{dd,HD}}_g)^2}{d_g} = (s^9+t^9+u^9)  \frac{[12][34]}{\langle 12\rangle \langle 34\rangle}\,.
\end{equation}
In general, any four-point symmetric vector numerator \cite{Carrasco:2022jxn} can be constructed from two $D$-dimensional gauge theory building blocks,
\begin{equation}
n_s^{\text{dd,vec,1}} = f_{12}f_{34}\,, \qquad n_s^{\text{dd,vec,2}} = f_{1324}\,,
\end{equation}
by composing with the symmetric scalar numerators, $n_s^{\text{dd,1}} = s$ and  $n_s^{\text{dd,2}} = tu$. For example, with these building blocks, symmetric double-copy construction captures the exceptional four graviton amplitude of Ref.~\cite{Boels:2016xhc}, which only considered linear combinations of gauge theory amplitudes rather than gauge theory numerators. We note that a similar construction was recently identified in Ref.~\cite{Bourjaily:2023tcc} beyond four-point, which demonstrated that gravity amplitudes permit double copy construction in terms of gauge-invariant numerators that do not obey any algebraic relations between themselves, much like the symmetric numerators above in \eqn{eq:symNum2loop}. We see further exploring higher multiplicity examples of symmetric double copy as an important direction of future study.

@@

[\mathcal{H}^{(ijkl)}] = 1,0,1,1,1,1,2,1,2,2,2,2,3,2,3,3,3,\cdots\,.

@@

[\mathcal{H}^{(ij)(kl)}] = 1,1,2,2,3,3,4,4,5,5,6,6,7,7,8,8,\cdots\,.

@@

Since each of these Hilbert series has an iterative structure where they compound at successively higher orders, they can be easily derived. Take for example the Hilbert series for symmetric invariants, which grows periodically at $\mathcal{O}(\alpha^{6n})$. To derive this series, we can begin with the geometric series,

@@

\left[\frac{1}{1-\alpha} \right]= 1,1,1,1,1,1,1,1,\cdots\,.

@@

\left[\frac{1-\alpha+\alpha^2}{1-\alpha} \right]= 1,0,1,1,1,1,1,1,\cdots\,.

@@

\mathcal{H}^{(ijkl)} =  \frac{1-\alpha+\alpha^2}{1-\alpha} (1+\alpha^{6}+\alpha^{12}+\alpha^{18}+\cdots )\,.

@@

\mathcal{H}^{(ijkl)} = \frac{1-\alpha+\alpha^2}{1-\alpha} \frac{1}{1-\alpha^6} =  \frac{1}{(\alpha-1)^2(\alpha+1)(\alpha^2+\alpha+1)} \, .

@@

\mathcal{T}^{F^2F^2}_{(x,y)} \sim s_{ij}^x(s_{ik}s_{jk})^y\,, \qquad \mathcal{T}^{F^4}_{(x,y)} \sim s_{ij}^x(s_{ik}s_{jk})^y \,, \qquad  \mathcal{T}^{F^3}_{(x,y)} \sim \sigma_3^x\sigma_2^y\,,

@@

\mathcal{T}^{(--++)}_{(x,y)} \sim s_{ij}^x(s_{ik}s_{jk})^y \,, \qquad \mathcal{T}^{(-+++)}_{(x,y)} \sim \sigma_3^x\sigma_2^y \,, \qquad \mathcal{T}^{(++++)}_{(x,y)} \sim \sigma_3^x\sigma_2^y\,.
